### 2025-05-08 Chapters/discussion.tex
changes

\section{Discussion}

@@

\subsection{Limitations and Future Directions}
Despite its strengths, the proposed method has notable limitations. The combined GCN-Transformer model incurs higher computational costs compared to lighter architectures like CNNs or LSTMs, which may restrict deployment in resource-constrained environments. The Sleep-EDF dataset, although widely used, includes a limited sample of subjects, potentially affecting generalizability. Furthermore, while acceleration via Apple M1 and Metal APIs improves efficiency, larger models would benefit from more powerful hardware such as GPUs or TPUs. Future work may explore more scalable architectures and validate the approach on larger, more diverse datasets.

\section{Conclusion}

The proposed SleepGCN-Transformer architecture demonstrates superior performance in sleep stage classification, achieving 93.12\% training accuracy and 93.04\% validation accuracy. By integrating GCNs to model spatial interdependencies and Transformers to extract temporal patterns, the model effectively captures the complexity of sleep physiology. The application of Focal Loss contributes to robust handling of class imbalance, improving classification across all stages. Feature importance analysis indicates that EMG and the EEG Pz-Oz channel play a critical role in prediction. This approach not only improves predictive performance but also enhances interpretability, laying a foundation for future work in explainable and clinically applicable sleep analysis tools.

>>>>

@@

\subsection{Limitations and Future Directions}
Despite its strengths, the proposed method has notable limitations. The combined GCN-Transformer model incurs higher computational costs compared to lighter architectures like CNNs or LSTMs, which may restrict deployment in resource-constrained environments. The Sleep-EDF dataset, although widely used, includes a limited sample of subjects, potentially affecting generalizability. Furthermore, while acceleration via Apple M1 and Metal APIs improves efficiency, larger models would benefit from more powerful hardware such as GPUs or TPUs.

### 2025-05-08 Title/conclusion.tex
changes

Our proposed SleepGCN-Transformer model achieves 93.12\% training accuracy and
93.04\% validation accuracy, demonstrating its effectiveness in sleep stage classification.
The integration of Graph Convolution Networks (GCN) captures spatial dependencies
across EEG, EOG, and EMG channels, while the Transformer extracts temporal patterns. The
use of Focal Loss enhances class balancing, improving performance on underrepresented sleep
stages. Feature importance analysis highlights EMG and EEG Pz-Oz as key predictors. This
robust approach lays the foundation for future work in Explainable AI, enabling medical
professionals to interpret AI-driven sleep diagnostics effectively.

>>>>

The proposed SleepGCN-Transformer architecture demonstrates superior performance in sleep stage classification, achieving 93.12\% training accuracy and 93.04\% validation accuracy. By integrating GCNs to model spatial interdependencies and Transformers to extract temporal patterns, the model effectively captures the complexity of sleep physiology. The application of Focal Loss contributes to robust handling of class imbalance, improving classification across all stages. Feature importance analysis indicates that EMG and the EEG Pz-Oz channel play a critical role in prediction. This approach not only improves predictive performance but also enhances interpretability, laying a foundation for future work in explainable and clinically applicable sleep analysis tools.
